---
title: "Simulation pipeline"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

This script is the repeatable script for measuring changes in the different diversity metrics and how they are affected by the different stressors. 

## Requirements for the simulations

In order to run the simulations (on your machine, cluster, etc...) you'll need to load the following in your current environment:

```{r, eval = FALSE}
## Getting the non-CRAN (yet) libraries
library(devtools)
devtools::install_github("TGuillerme/dads")
```

```{r}
## Libraries
require(dads)
require(dispRity)
require(BAT)
require(FD)
require(ks)
require(TPD)
```

> Note that the libraries do not need to be loaded (i.e. no need for `library()`) but needs to exist.

You'll also need the following custom functions (located in [here on github](https://github.com/TGuillerme/eco_metrics_simulations/tree/master/Functions))

```{r}
## Loading the functions
source("../Functions/melodic.rao.R")
source("../Functions/optim.replicate.R")
source("../Functions/make.simulation.pipeline.R")
source("../Functions/simulation.pipeline.R")
```




## Setting up the simulations


## Subtitle


```{r, eval = FALSE, message = FALSE, fig.width = 12, fig.height = 12, echo = FALSE}
```
